\Title

### 2023-06-27 2.REVISION/REPLY_REVIEWERS/reply_reviewers.tex

\documentclass[11pt,letter]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amstext,pdfsync}
\usepackage[dvips]{color}
\usepackage{tcolorbox}
\usepackage{mathpazo}
\usepackage{amssymb,xspace,enumerate}
\definecolor{LineColor}{rgb}{0.4,0.4,0.4}

%--------------------------------------------------%
\definecolor{shadecolor}{gray}{.9}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\newenvironment{myquote}{%
\bigskip%
\begin{tcolorbox}[]\footnotesize}{%
\end{tcolorbox}%			
\bigskip}

%\headsep 0 true cm
%\textwidth 15.0 true cm
%\textheight 24.0 true cm
%\topmargin 0 true cm
%\oddsidemargin 0 true cm
%\evensidemargin 0 true cm
%\headsep 0 true cm
\usepackage{geometry}

%\setlength{\parindent}{2ex}

\newcommand*{\vhc}{\textsc{vhc}\xspace}
\newcommand*{\vhcs}{\textsc{vhc}s\xspace}
\newcommand*{\vcg}{\textsc{vcg}\xspace}
\newcommand*{\vcgs}{\textsc{vcg}s\xspace}
\newcommand*{\R}{\mathbb{R}}
\newcommand*{\Q}{\mathcal{Q}}
\newcommand*{\X}{\mathcal{X}}
\newcommand*{\specone}{\textsc{spec 1}\xspace}
\newcommand*{\spectwo}{\textsc{spec 2}\xspace}
\newcommand{\NC}{\textsc{nlc}\xspace}
\DeclareMathOperator{\Image}{Im}

\newcommand{\drawline}
{   \hspace*{-\marginparsep minus \marginparwidth}
	\begin{minipage}[t]{\textwidth+\marginparwidth+\marginparsep}%
		{\large \bfseries {} \hfill {}}\\[-0.15\baselineskip]%
		\textcolor{LineColor}{\rule{\columnwidth}{2 pt}}%}
	\end{minipage}}

\newcommand{\myver}{\vspace{0.5cm}}

\newcommand{\mybox}[1]{{\color{blue}\fbox{#1}}}

%\include{Bibtex_DistTeleop}

\title{\vspace{-1cm} {\Large \bf \color{darkblue}{Reply to the Associate
      Editor and Reviewers}\\[0.6cm] \large {\bf From gymnastics to virtual nonholonomic constraints: energy
injection, dissipation, and regulation for the acrobot
      }\\[0.2cm] {(Revision of  IEEE TCST 22-0089)\\ A.~Moran-MacDonald, M.~Maggiore, X.~Wang}}} \author{}
\date{\today}

\begin{document}
\vspace{-4cm}
\maketitle

%\drawline
\pagenumbering{gobble}

\indent We sincerely thank the Associate Editor for collecting two thoughtful reviews and for the 
constructive remarks on this submission. We have revised the paper addressing the reviewers'
comments. Below are detailed replies to the Associate Editor and the reviewers.
\pagenumbering{arabic}
%---------------------------------------------------------------------------%
%\vspace{-1cm}
\subsection*{\textcolor{darkblue}{Reply to the Associate Editor}} 
\vspace{0cm}

\begin{myquote}
We have received only two reviews for this paper, which are - however -
sufficiently informative for a recommendation  to be  issued. Overall,
the reviewers praise the high quality of the paper, but request that a
sharper focus is placed on its contribution with respect not only on
the literature on virtual nonholonomic constraints, but the  landscape
of energy-based nonlinear control. While this may appear a daunting
task, the authors are invited to put their work into perspective as
suggested by reviewer no.2.  The remaining issues appear to be well
within the authors' reach.  
\end{myquote}

In this revision, we have significantly expanded the introduction to include a \emph{detailed} overview of the literature on orbital stabilization mentioned by reviewer 2, contrasting our ideas to existing ones. We have also included a new subsection titled ``Contributions of this article'' in which we better showcase the main results of the paper.  We have also addressed all other issues raised by the two reviewers.

\newpage
%---------------------------------------------------------------%
\vspace{0.2cm}
\subsection*{\textcolor{darkblue}{Reply to Reviewer 2}} 
\vspace{0.2cm}

We wish to sincerely thank the reviewer for the constructive comments,
which we have  addressed in this revision. Below are
detailed replies to the reviewer's remarks.

\begin{myquote}
The problem of designing a feedback control for the two-link acrobot
system through virtual nonholonomic constraints is addressed in this
paper. \((\cdots)\) 
The paper is technically correct and the proofs of Theorem 4 and
Theorem 5 are mathematically rigorous and consistent with the
statements.
\end{myquote}

We thank the reviewer for this assessment.

\begin{myquote}
However, my major concern is that neither the benefit nor
the novelty of considering stabilization by assigning virtual
nonholonomic constraints are clearly articulated. The literature on how
to design nonlinear controllers, possibly passive, for achieving either
swing-up and orbital stabilization or swing-up and asymptotic
stabilization is extensive. In terms of the acrobot, most of the
existing approaches rely on the construction of Casimir functions
(conserved quantities) to help shape the energy of the closed-loop
system via Control-by-Interconnection (CbI) as in [R9], [R8], the
potential energy shaping of the closed-loop system via Interconnection
and Damping Assignment (IDA-PBC) as in [R1], [R2], [R3], [R7], orbital
stabilization via Immersion and Invariance (I\&I) as in [R4], [R5], and
finally orbital stabilization via energy shaping and with the injection
of both pumping and damping as in [R6]. As an application-based
research paper, I would have expected an application of established
research findings and a comparison with other existing approaches. In
addition to showing the effectiveness of implementing virtual
nonholonomic constraints, it would be valuable to discuss the benefit
of the method when compared with the approaches mentioned beforehand. 
\end{myquote}

The reviewer makes a good point, and we completely agree that not placing our paper in a wider context was an omission. This revision \emph{significantly expands} the literature treatment and placed our work in a wider context. The introduction has been significantly expanded with a new subsection titled ``Related literature'' in which we review other techniques for orbital stabilization: I\&I, IDA, VHC-based, and the Fantoni-Lozano approach. We also stress the differences between these approaches and the ideas of this submission.

\begin{myquote}
Furthermore, the following issues should be addressed:

1.	Please define q before the Hamiltonian function given in by the
equations (6). I assume that the generalized coordinate q is composed
of (qu, qa), however, the structure of q is not defined anywhere in the
manuscript.
\end{myquote}

Done, and thank you for raising this point.

\begin{myquote}
 2.	Please enumerate the Hamiltonian function H(q,p) in (6) with
(6a) and the dynamics associated with H(q, p) with (6b). It is quite
uncommon to write "The dynamics of the acrobot in (q,p) coordinates are
given in (6)î. I would suggest, instead, writing îThe dynamics of the
acrobot with Hamiltonian function (6a) are given in (6b)î.  
\end{myquote}

Done, and thank you.

\begin{myquote}
3.	The manuscript deals with the acrobot model, which yields a
Hamiltonian system described by an inhomogeneous differential equation
with a constant input matrix. What is the benefit of introducing the
general Hamiltonian system as in (8) with a nonconstant input matrix B
and then assuming B constant? For ease of presentation, I would suggest
considering directly the inhomogeneous Hamiltonian system used in the
paper.  
\end{myquote}

Done. As suggested by the reviewer, we now directly consider a model with \(B\) constant.

\begin{myquote}
4.	If Lemma 1 and Theorem 1 are recalled from another paper,
please cite the paper which contains the statements and proofs, and
remove the proofs from this paper. Otherwise, if Lemma 1 and Theorem 1
are original results or enhance some other previous findings, then keep
the proofs but emphasise the originality of the results.  
\end{myquote}

We did not take Lemma 1 and Theorem 1 from the literature. However,
the results are relatively obvious and for this reason we do not
showcase them. We clarified this point in a footnote.

\begin{myquote}
5.	Please restate Definition 2. The expression îthe output e =
h(q,p) is of relative degree 2, 2, ..., 2î is unusual. The output is
not of relative degree, rather the system associated with the output e
= h(q,p) has a relative degree. In addition, relative degree 2, 2, ...,
2 is unclear. I think the authors wanted to say that the system has
relative degree 2 for each e\_i.
\end{myquote}

Fixed, thank you. The correct term is \emph{vector relative degree}. We also added a reference to the book by Isidori where this notion is defined.

\begin{myquote}
6.	From Definition 2 results that, if you define the output as e =
h(q,p), where h satisfies the relation h(q,p) = 0, you are looking for
a special characterization of the zero dynamics of your closed-loop
system. This approach reminds me of the construction of the invariant
manifold used to construct a stabilizer through I\&I [R5]. Is there any
relation between the two?  
\end{myquote}

In this revision we have given a detailed comparison between the  I\&I technique and the approach proposed in this paper. Both approaches involve the stabilization of a submanifold of the state space on which one has ``desirable'' behaviour. While in the I\&I literature related to orbital stabilization one assumes that the dynamics on the submanifold are those of an oscillator, in this paper we show that the dynamics on the constraint manifold induced by the proposed VNHC are ``expansive'' or ``contractive'' in that orbits gain energy or lose energy, depending on the sign of a design parameter. In particular the dynamics on our constraint manifold do not contain closed orbits. 
\begin{myquote}
  7.	The output e = h(q,p), which is here assumed relative degree 2,
rules out the most common passive outputs because the relative degree
of the passive output is at most 1. Hence, the virtual nonholonomic
constraints seem to be weak compared to the IDA-PBC approach in [R6].
Please, clarify their relationship. Moreover, the original results in
Section IV.A also seems to have much in common with the
pumping-and-damping technique in [R6].
\end{myquote}

Having relative degree two rather than one does not imply any weakness per se. There is indeed a philosophical  similarity between pumping and damping and the ideas of this submission. This is clarified in the revised introduction. Despite the philosophical similarity, the two techniques are very different and we point this out.

\begin{myquote}
8.	In the sentence before equation (14), what type of stability do
you refer to? 
\end{myquote}

We were referring to asymptotic stability of the constraint manifold. We have modified the language to make this clear.

\begin{myquote}
9.	As I have already mentioned, the subsection îComparison with
existing literatureî only discusses previous work on virtual
nonholonomic constraints and a comparison with other consolidated
methods such as I\&I, CbI, and IDA-PBC is missing. Moreover, a
comparison should be given also in terms of simulations showing the
effectiveness of the virtual nonholonomic constraints when compared
with the existing methods.   
\end{myquote}

We have addressed this point in great detail. We have not however included a simulation comparison because for it to be meaningful, the comparison would require  detailed discussions and an overview of the various techniques  which are beyond the scope of this paper. We hope that the detailed comparison provided in the introduction will at least partially address the reviewer's concern.

\begin{myquote}
10.	Some general comments: the introduction should be improved by
motivating and properly discussing the main results achieved within the
paper; it would be preferable to conclude the introduction with the
organization of the paper; the reading of the article would be easier
if the problem formulation is presented after the preliminary section.  
\end{myquote}

We agree. We have added two new subsections titled ``Contributions of this paper'' and ``Paper organization'' in which we do exactly what the reviewer recommended.

\newpage
%---------------------------------------------------------------------------------%
\vspace{0.2cm}
\subsection*{\textcolor{darkblue}{Reply to Reviewer 4}}
\vspace{0.2cm}

We wish to sincerely thank the reviewer for the constructive comments. Below are
detailed replies to the reviewer's remarks.

\begin{myquote}
My main concern: The topic of virtual (nonlinear) nonholonomic
constraints has its roots in the works of Grizzle and Griffin in CDC
2015. Since then that notion was used and employed in a very
interesting context as the one in the present paper. Nevertheless, a
mathematical foundation of the subject as happens with virtual
holonomic constraints (done by the second author of the paper under
review) was not provided yet. Some effort has been done recently to
formulate virtual nonholonomic constraints from a geometric framework
in the linear (https://arxiv.org/abs/2207.01299) and affine cases
(https://arxiv.org/abs/2301.03890) that I guess they merit to be
mentioned in the literature review of the paper.
\end{myquote}

Thank you very much for flagging this paper. We are now citing it and discussing it both in the introduction and in Section IIIB on virtual nonholonomic constraints.

\begin{myquote}
  My recommendation is based on the next aspects of the paper:

(1)The paper is mathematically correct and self-contained.

(2)The paper is accessible to a wide range of researchers in control
and robotics

(3)The paper demonstrates the theory with real-life experimental.

(4) The experiments are understandable for a wide range of researchers
in control and robotics

(5) The paper is original and the topic is on current research.

I recommend acceptance of the manuscript.
\end{myquote}

We sincerely appreciate the positive assessment of our work.
\end{document}

### 2023-06-27 2.REVISION/main.tex
Revised paper, resubmitted to the IEEE TCST. Also two folders containing literature and the reply to reviewers.

\usepackage{generic}
\usepackage{hyperref}
%/========== Preamble ==========/%
% Packages required by ieeecolor
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{xspace}
\usepackage{enumerate}
%---------- More packages ----------%
% NOTE: cannot use amsthm for some reason
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}

@@

@@

%/========== Main Document ==========/%
\begin{document}
\title{\Title}
\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
\IEEEmembership{Member, IEEE}*, and Xingbo Wang
\thanks{Manuscript submitted for review on \today.}
\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca),
M. Maggiore (e-mail: maggiore@control.utoronto.ca), and
X. Wang (e-mail: xingbo.wang@utoronto.ca) are with the Department of
Electrical and Computer Engineering, University of Toronto, ON, Canada.}
} %/author

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.40\linewidth}
        \includegraphics[width=\linewidth]{acrobot_gymnast.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.48\linewidth}
        \includegraphics[width=\linewidth]{simple_acrobot_model.pdf}
    \end{subfigure}
    \caption{A simplified two-link acrobot as a model for a gymnast.
    Image modified from \cite{xingbo_thesis}.}
    \label{fig:acrobot}
\end{figure}

Previous attempts at acrobot giant generation have involved
trajectory tracking, partial feedback linearization, or other energy-based
methods
( see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While all these approaches succeed at making the acrobot rotate around the
bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
That is, none of these leg controllers track a function of the acrobot's body
angle and velocity.
In 2016, Wang designed a controller which tracked the
body angle and an \textit{estimate} of the velocity, but not the velocity itself
\cite{xingbo_thesis}.
His approach was a preliminary version of a recent technique know as the method
of virtual nonholonomic constraints.

Virtual nonholonomic constraints (\vnhcs) have been used for human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and they have shown marked improvements to the field of bipedal locomotion 
\cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Indeed, they produce more robust walking motion in biped robots than
other virtual constraints which do not depend on velocity
\cite{nhvc_incline_walking}.
In particular, \vnhcs may be capable of injecting and
dissipating energy from a system in a robust manner, all while producing
realistic biological motion.

In this article, we design a virtual nonholonomic constraint which provably
injects energy into the acrobot through human-like giant motion.
%In particular, any acrobot constrained by this \vnhc is guaranteed to perform
%rotations around the bar.
%Under suitable conditions on the acrobot's physical parameters, this \vnhc
%also enables the acrobot to rotate around the bar at a desired speed.
We will provide simulations to show how one can use \vnhcs to regulate energy,
along with experimental results which demonstrate the robustness of this
behaviour to model uncertainty, sensor noise, and a variety of external
disturbances.

@@

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
We will use the simplified acrobot model in Figure
\ref{fig:acrobot}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
The acrobot's configuration is described in generalized coordinates
\((q_u,q_a)\) on the configuration manifold 
\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
\(q_a\) is actuated.
We ignore dissipative forces in this model.

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear later in this article, we
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.

@@

%/========== \vnhc ==========/%
\section{Preliminaries on \vnhcs}\label{sec:vnhc}
Before embarking on the design problem, we must summarize the relevant
theory of virtual nonholonomic constraints for a class of mechanical systems we
call ``simply actuated Hamiltonian systems". 
The results we provide in Section \ref{sec:vnhc-vnhc} are not novel:
they are a special case of the results in
\cite{hybrid_zero_dynamics_bipedal_nhvcs}.

\subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
Take a mechanical system modelled with generalized coordinates 
\(q = (q_1, \ldots, q_n)\) on a configuration manifold
\(\mathcal{Q} = \Rt{T_1} \times \cdots \times \Rt{T_n}\), where
\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
displacement. The corresponding generalized velocities are 
\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).

>>>>

\usepackage{generic}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=red,
    filecolor=black,
    linkcolor=blue,
    urlcolor=black
  }
% /========== Preamble ==========/%
% Packages required by ieeecolor
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{xspace}
\usepackage{enumerate}
\usepackage{xcolor}
%---------- More packages ----------%
% NOTE: cannot use amsthm for some reason
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}

@@

\definecolor{MM-blue}{HTML}{2883E6}
\newenvironment{MM}[1]{{\color{MM-blue}#1}}{}

@@

%/========== Main Document ==========/%
\begin{document}
\title{\Title}
\author{Adan Moran-MacDonald, \IEEEmembership{Member, IEEE}, Manfredi Maggiore,
\IEEEmembership{Senior Member, IEEE}*, and Xingbo Wang
\thanks{Manuscript submitted for review on \today.}
\thanks{A. Moran-MacDonald (e-mail: adan.moran@mail.utoronto.ca),
M. Maggiore (e-mail: maggiore@control.utoronto.ca), and
X. Wang (e-mail: xingbo.wang@utoronto.ca) are with the Department of
Electrical and Computer Engineering, University of Toronto, ON, Canada.}
} %/author

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.40\linewidth}
        \includegraphics[width=\linewidth]{FIGURES/acrobot_gymnast.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.48\linewidth}
        \includegraphics[width=\linewidth]{FIGURES/simple_acrobot_model.pdf}
    \end{subfigure}
    \caption{A simplified two-link acrobot as a model for a gymnast.
    Image modified from \cite{xingbo_thesis}.}
    \label{fig:acrobot}
\end{figure}

Previous attempts at acrobot giant generation have involved
trajectory tracking, partial feedback linearization, or other energy-based
methods
(see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While all these approaches succeed at making the acrobot rotate around the
bar, none of them use the results of \cite{pendulum_length_giant_gymnastics}.
That is, none of these leg controllers track a function of the acrobot's body
angle and velocity.
In 2016, Wang designed a controller which tracked the
body angle and an \textit{estimate} of the velocity, but not the velocity itself
\cite{xingbo_thesis}.
His approach was a preliminary version of a recent technique know as the method
of virtual nonholonomic constraints.

Virtual nonholonomic constraints (\vnhcs) have been used for human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and they have shown marked improvements to the field of bipedal locomotion 
\cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Indeed, they produce more robust walking motion in biped robots than
other virtual constraints which do not depend on velocity
\cite{nhvc_incline_walking}.
In particular, \vnhcs may be capable of injecting and
dissipating energy from a system in a robust manner, all while producing
realistic biological motion. \MM{The recent paper~\cite{SimStraBloCol23} initiates a systematic geometric investigation of \vnhcs that are affine functions of the configuration velocities.}

\MM{\textbf{Contributions of this article:}}
In this article, we design a \vnhc which provably
injects \MM{or dissipates} energy into the acrobot \MM{depending on the sign of a design parameter. The proposed constraint produces a} human-like giant motion.
\MM{The main result of the paper is presented in Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}.   Theorem~\ref{thm:acrobot-oscillations} shows that once the virtual constraint has been enforced, for any nonzero ``low-energy'' initial condition, the proposed controller injects energy in the acrobot making it exit, in finite time, any compact subset of a certain region of the state space where the acrobot oscillates about its pivot point without performing full rotations. Theorem~\ref{thm:acrobot-rotations} handles the case when the acrobot is initialized in the complement of the above region, and it shows that under a mild assumption on the physical parameters, the acrobot  is guaranteed to keep increasing its energy while performing full rotations about the pivot point. By changing the sign of the design parameter, the results of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} hold for energy dissipation. By combining these energy injection/damping properties, we then propose an elementary switching logic to attain approximate energy regulation. The results of this paper are demonstrated both in simulation and experimentally. The experimental results, in particular,} demonstrate the robustness of \MM{the controller against} model uncertainty, sensor noise, and a variety of external
disturbances.

\MM{\textbf{Related literature:} While the focus of this paper is on energy injection and dissipation, a closely related concept is that of orbital stabilization to a particular energy level set. The  problem of orbital stabilization has been investigated extensively and we mention four main approaches.

  One approach, applicable chiefly to underactuated mechanical systems with degree of underactuation one, relies on virtual holonomic constraints to create a target closed orbit and some mechanism to stabilize this orbit. In~\cite{ShiPerWit05}, the mechanical system is linearized around the orbit and a linear time-varying controller is designed to stabilize it. In~\cite{dynamic_vhcs_stabilize_closed_orbits}, the constraint is embedded in a one-parameter family, and a dynamic controller is designed to adapt the parameter so as to stabilize the target orbit.

  A second approach to orbital stabilization, developed in~\cite{OrtYiRomAst20} and applicable to general control-affine systems, is an adaptation of the immersion and invariance (I\&I) technique of~\cite{AstOrt03}. In this approach, one seeks a controlled invariant submanifold of the state space on which the system has desired oscillator dynamics, and designs a feedback controller to stabilize this submanifold. The result is a controller that stabilizes \emph{some} closed orbit on the manifold, depending on the initial condition. This is in contrast to the approaches in~\cite{ShiPerWit05,dynamic_vhcs_stabilize_closed_orbits} that are able to fix the closed orbit to be stabilized.

  A third approach, developed in~\cite{YiOrtWuZha20}, is an adaptation of the passivity-based interconnection and damping assignment (IDA) technique of~\cite{OrtSpoGomBlan02} and is applicable to underactuated port-Hamiltonian systems. In~\cite{YiOrtWuZha20}, the authors assume that the target closed orbit is contained in a two-dimensional plane and present two ways of stabilizing it. The first way relies on a modification of energy shaping via IDA in which the component of the shaped energy on the plane has a minimum at the target closed orbit so that its graph resembles a Mexican sombrero.  The second way represents the closed orbit as an energy level set and assigns the damping matrix in such a way that the controller pumps  energy when the energy is below the target level and damps energy otherwise.

  A final approach, developed in~\cite{FanLoz02}, stabilizes homoclinic orbits of pendulum-like systems using energy methods and Lyapunov-based passivity techniques. The stabilization of a homoclinic orbit allows the designer to inject energy in the mechanical system, and it typically yields large domains of attraction.

 We now place the contributions of this paper in the context of the literature just reviewed. It is important to bear in mind that the ideas put forth here are limited in scope to the acrobot. They are nonetheless notable in that they explore a mechanism to pump or dissipate energy  that is guaranteed to work in a large region of the state space. Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} make this precise. In contrast, many the approaches found in the literature and reviewed  above yield only local asymptotic stabilization of a closed orbit. When it comes to the acrobot example, the employment of \vnhcs allows one to incorporate knowledge about giant maneuvers in gymnastic into the control design, something that cannot be easily achieved with existing techniques.

 There are similarities between the virtual constraint approach and the I\&I technique for orbital stabilization  proposed in~\cite{OrtYiRomAst20} in that both approaches seeks a controlled invariant submanifold on which the closed-loop system exhibit desired dynamics. But while in~\cite{OrtYiRomAst20} one wishes to obtain oscillator dynamics on the submanifold, in this paper we wish to obtain expansive or contractive dynamics (depending on the sign of a design parameter) which rules out oscillator dynamics.

In this paper, like in~\cite{YiOrtWuZha20}, we design a controller for energy pumping/damping, but the energy pumping/dumping mechanism proposed here does not arise from any form of energy shaping. In particular, it does not arise from an assignment of the damping matrix of the closed-loop system.

Finally, while in~\cite{FanLoz02} the emphasis is typically the stabilization of a specific energy level set corresponding to the homoclinic orbit of an unstable equilibrium, the technique used in this paper to inject energy does not involve the stabilization of any homoclinic orbit, and as such it does not limit the energy level that can be attained.
  }

  \MM{\textbf{Paper organization:}
In Section~\ref{sec:problem-formulation} we review the acrobot model and formulate the problem investigated in this paper. Section~\ref{sec:vnhc} presents a class of simply actuated Hamiltonian systems and preliminary notions on \vnhcs. In particular, Theorem~\ref{thm:vnhc-regularity} in this section characterizes the constrained dynamics arising from a class of \vnhcs. This result is then compared with existing literature on the subject. Section~\ref{sec:acrobot} presents the proposed \vnhc for the acrobot and the two main results of the paper, Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations} mentioned earlier. Sections~\ref{sec:simulations} and~\ref{sec:experiments} present simulations and experimental results. Section~\ref{sec:proof} presents the proof of Theorems~\ref{thm:acrobot-oscillations} and~\ref{thm:acrobot-rotations}, and Section~\ref{sec:conclusion} presents concluding remarks.
}

@@

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
We will use the simplified acrobot model in Figure
\ref{fig:acrobot}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
The acrobot's configuration is described in generalized coordinates
\MM{\(q=(q_u,q_a)\)} on the configuration manifold 
\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
\(q_a\) is actuated.
We ignore dissipative forces in this model.

\MM{In what follows, we denote by \(c_u \eqdef \cos(q_u)\), \(c_a \eqdef \cos(q_a)\), and 
\(c_{ua} \eqdef \cos(q_u + q_a)\); likewise, \(s_u \eqdef \sin(q_u)\), 
\(s_a \eqdef \sin(q_a)\), and \(s_{ua} \eqdef \sin(q_u + q_a)\).
}

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2 c_a\right) & 
        ml^2\left(1+ c_a \right) \\
        ml^2\left(1+ c_a\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2 c_u + c_{ua}\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear later in this article, we
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
\MM{To this end, recalling the partitioning  \(q= (q_u,q_a)\), we define the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\) and the Hamiltonian function
\begin{equation}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) = \frac{1}{2}p\tpose \Minv(q) p -
   mgl\left(2 c_u + c_{ua}\right).
  \end{equation}
The dynamics of the
acrobot with Hamiltonian function~\eqref{eqn:acrobot-hamiltonian} are
\begin{equation}
  \label{eq:acrobot:model}
  \begin{aligned}
&    \dot{q} = \Minv(q) p,\\
&    \dot{p}_u = -mgl\left(2s_u + s_{ua}\right),\\
&     \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p - m g l s_{ua} + \tau,
  \end{aligned}
\end{equation}
where \(\tau\), the control input, is the hip joint torque affecting only the dynamics of
\(p_a\).}

@@

%/========== \vnhc ==========/%
\section{Preliminaries on \vnhcs}\label{sec:vnhc}
Before embarking on the design problem, we must summarize the relevant
theory of \vnhcs for a class of mechanical systems we
call ``simply actuated Hamiltonian systems". 
\MM{Although the results in Section \ref{sec:vnhc-vnhc} do not appear in the literature in the same form, we will show that} they are a special case of results that appeared in
\cite{hybrid_zero_dynamics_bipedal_nhvcs}.

\subsection{Simply Actuated Hamiltonian Systems} \label{sec:vnhc-sah}
\MM{Consider} a mechanical system modelled with generalized coordinates 
\(q = (q_1, \ldots, q_n)\) on a configuration manifold
\(\mathcal{Q} = \Rt{T_1} \times \cdots \times \Rt{T_n}\), where
\(T_i = 2\pi\) if \(q_i\) is an angle and \(T_i = \infty\) if \(q_i\) is a
displacement. The corresponding generalized velocities are 
\(\dot{q} = (\dot{q}_1,\ldots,\dot{q}_n) \in \R^n\).

Suppose this system has Lagrangian
\(\mathcal{L}(q,\dot{q}) = 1/2~\dot{q}^T D(q) \dot{q} - P(q)\),
where the potential energy 
\(P : \mathcal{Q} \rightarrow \mathbb{R}\) 
is smooth, and the inertia matrix 
\(D : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
is smooth and positive definite for all \(q \in \mathcal{Q}\).
The \textit{conjugate momentum} to \(q\) is the vector
\(p \eqdef \partial\mathcal{L}/\partial\dot{q} = D(q) \dot{q} \in \R^n\).
As per \cite{landau_mechanics}, 
the \textit{Hamiltonian} of the system in \((q,p)\) coordinates
is
\begin{equation}\label{eqn:hamiltonian}
    \mathcal{H}(q,p) = \frac{1}{2} p\tpose D\inv(q) p + P(q)
    ,
\end{equation}
with dynamics
\begin{equation}\label{eqn:hamiltonian-eom-general}
    \begin{cases}
        \dot{q} = \nabla_p\mathcal{H} 
        , \\
        \dot{p} = -\nabla_q\mathcal{H} + \MM{B} \, \tau
        ,
    \end{cases}
\end{equation}
where \(\tau \in \R^k\) is a vector of generalized input forces and the input
matrix \MM{\(B \in \R^{n \times k}\) is assumed to be constant and of full rank \(k\).}
If \(k < n\), we say the system is \textit{underactuated} with degree of
underactuation \((n-k)\).

Using the matrix Kronecker product, it is easy to show that
\eqref{eqn:hamiltonian-eom-general} expands to
\begin{equation*}\label{eqn:hamiltonian-full-dynamics}
     \begin{cases}
        \dot{q} = D\inv(q)p 
        , \\
        \dot{p} = -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
        - \nabla_q P(q) + \MM{B} \,\tau
        . \\
    \end{cases}
\end{equation*}
%\begin{align*}
%    \dot{q} &= D\inv(q)p 
%    ,\\
%    \dot{p} &= -\frac{1}{2} (\Id{n} \otimes p\tpose) \nabla_q D\inv(q) p
%        - \nabla_q P(q) + B(q) \tau
%    .
%\end{align*}

Because \(\tau\) is transformed by \(\MM{B}\), it is not obvious how any
particular input force \(\tau_i\) affects the system.
% As a first step in addressing this problem, we make the following assumptions.
